---
title: "amoham57_Assignment5"
author: "Abdul Sahil Mohammed"
date: "2024-04-08"
output: html_document
---
```{r}
getwd()
```

# Necessary libraries
```{r}
library(stats)
library(cluster)
library(factoextra)
library(dplyr)
library(caret)
library(tidyr)
library(fpc)
```
 
```{r}

library(readxl)
creals.Data <- read.csv("C:/Users/abdul/OneDrive/AssignmentFML/Cereals.csv")
ext.cereals <- read.csv("C:/Users/abdul/OneDrive/AssignmentFML/Cereals.csv", row.names = 1)
```

# Eliminating the columns from the data frame
```{r}
ext.cereals <- subset(ext.cereals, select = -c(mfr, type))
```
# Show the initial few rows of the updated dataset.
```{r}
head(ext.cereals)
str(ext.cereals)
```
# leaving out missing values
```{r}
ext.cereals <- na.omit(ext.cereals)
head(ext.cereals)
```
# Make the numerical data normal.
```{r}
cerl.numeric <- ext.cereals %>% select_if(is.numeric)
cerl.scaled <- scale(cerl.numeric)
head(cerl.scaled)
```
## Agnes Implementation
```{r}
ext.cereals.E <- dist(ext.cereals, method = "euclidean")   
hc.1 <- hclust(ext.cereals.E, method = "complete")

```
# Plot Dendogram
```{r}
plot(hc.1, cex = 0.6, hang = -1)
```
# Computing Agnes using several other methods
```{r}
hc.single.crl <- agnes(cerl.scaled, method = "single")
hc.complete.crl <- agnes(cerl.scaled, method = "complete")
hc.average.crl <- agnes(cerl.scaled, method = "average")
hc.ward.crl <- agnes(cerl.scaled, method = "ward")
```

# Comparing the Agglomerative coefficients
```{r}
print(hc.single.crl$ac)
print(hc.complete.crl$ac)
print(hc.average.crl$ac)
print(hc.ward.crl$ac)
pltree(hc.ward.crl, cex = 0.6, hang = -1, main = "Dendrogram of the agnes")
```
# Altering the Dendrogram
```{r}
summary(ext.cereals)
```
# Plot/Dendrogram Reduction
# Reducing
```{r}
ext.cereals.E <- dist(ext.cereals, method = "euclidean") 
hc.ward.crl <- hclust(ext.cereals.E,method = "complete")
```
# Ploting 
```{r}
plot(hc.ward.crl, cex = 0.6)
rect.hclust(hc.ward.crl, k = 4, border = 1:4)
```
## Assessing whether using kMeans to the problem is appropriate and estimating its performance.
```{r}
set.seed(123)  
partition.clr <- sample(2, nrow(cerl.scaled), replace = TRUE, prob = c(0.7, 0.3))
cerls.A <- cerl.scaled[partition.clr == 1, ]
cerls.B <- cerl.scaled[partition.clr == 2, ]
```

# Hierarchical on A
```{r}
Hc.A <- agnes(cerls.A, method = "ward")

```

# Hierarchical on B
```{r}
Cluster.B.Assignment <- cutree(Hc.A, k = 3)
```

# Matching the lenth with A & B
```{r}
Cluster.B.Assignment <- Cluster.B.Assignment[partition.clr == 2] 
consistency <- table(Cluster.B.Assignment, partition.clr[partition.clr == 2])
print("Cluster Consistency Assignment: All.Bran.with.Additionsl.Fiber")
```
```{r}
hc.healthy <- agnes(cerl.scaled, method = "ward")
optimal.clusters <- which.max(hc.healthy$ac)
clustering.assignments <- cutree(hc.healthy, k = optimal.clusters)
desired.cluster.number <- 1
healthy.cerls <- ext.cereals[clustering.assignments == desired.cluster.number, ]
print("NutritiousCereals:Wheat")
```
```{r}
print(healthy.cerls)
```

