---
title: "amoham57_Assignment4"
author: "Abdul Sahil Mohammed"
date: "2024-03-17"
output: html_document
---
```{r}
library(factoextra)
library(ggplot2)
library(tidyverse)
library(ISLR)
library(NbClust)
library(cluster)
```
#Let's load the following DataSet 
```{r}
Pharmaceutical.data <- read.csv("C:/Users/abdul/OneDrive/AssignmentFML/Pharmaceuticals.csv")
```


```{r}
R <- na.omit(Pharmaceutical.data)
summary(R)
```
#Now lets Set row names of the data frame 'R' to the values in its first column
```{r}
row.names(R) <- R[,1]

Phar.data <- R[,3:11]

head(Phar.data)
```
#Here, we Scale the data in the 'Pharmacist.data' data frame to standardize variables
```{r}

Pharmacist.data <- scale (Phar.data)
#Display the rows of the 'Pharmacist.data' data frame
head(Pharmacist.data)
```

```{r}
fviz_nbclust(Phar.data, kmeans, method = "silhouette") + labs(subtitle = "silhouette Method")

```
# Set a seed for reproducibility
```{r}

set.seed(64060)

k8 <- kmeans(Phar.data, centers = 5, nstart = 26)

k8$centers
```

```{r}
fviz_cluster(k8, data = Pharmacist.data)
```
#here we Calculate the Euclidean distance matrix between observations in the 'Pharmacist.data' dataset
```{r}

distance <- dist(Pharmacist.data, method = "euclidean")
fviz_dist(distance)
```
```{r}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```

```{r}
fit <-kmeans(Pharmacist.data, 5)
aggregate(Pharmacist.data, by = list(fit$cluster), FUN=mean)
```
#Create a new data frame 'Pharmacist.data1' by combining the original data 'Pharmacist.data1' with the cluster assignments from 'fit$cluster'
```{r}
Pharmacist.data1 <- data.frame(Phar.data, fit$cluster)

# Display the contents of the newly created data frame 'Pharmacist.data1'
Pharmacist.data1
```
```{r}
clusplot(Pharmacist.data, fit$cluster, color = TRUE, shade = FALSE, labels = 2, lines = 0)
```




